---
title: 'Data Wrangling in \texttt{R}'
author: "Garrett Allen"
date: "Due Friday August 27, 5 PM EDT"
output: pdf_document
---
```{r setup}
library(tidyverse)
```
Today's agenda: Manipulating data objects; using the built-in functions, doing numerical calculations, and basic plots; reinforcing core probabilistic ideas.

***General instructions for homeworks***: Please follow the uploading file instructions according to the syllabus. You will give the commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file. Each answer must be supported by written statements as well as any code used. Your code must be completely reproducible and must compile. 

***Advice***: Start early on the homeworks and it is advised that you not wait until the day of. While the professor and the TA's check emails, they will be answered in the order they are received and last minute help will not be given unless we happen to be free.  

***Commenting code***
Code should be commented. See the Google style guide for questions regarding commenting or how to write 
code \url{https://google.github.io/styleguide/Rguide.xml}. No late homework's will be accepted.

***R Markdown Test***

0. Open a new R Markdown file; set the output to HTML mode and "Knit". This should produce a web page with the knitting procedure executing your code blocks. You can edit this new file to produce your homework submission.

***Working with data***

Total points on assignment: 10 (reproducibility) + 22 (Q1) + 9 (Q2) + 3 (Q3) = 44 points 

Reproducibility component: 10 points. 

1.  (22 points total, equally weighted) The data set **rnf6080.dat** records hourly rainfall at a certain location in Canada, every day from 1960 to 1980.

a. Load the data set into R and make it a data frame called \texttt{rain.df}.  What command did you use?
```{r read file}
rain.df <- read.delim("data/rnf6080.dat",sep = "",header = FALSE)
```
I used the read.delim command to read in the .dat file, with sep = "", since this was the format of the .dat file. I additionally had to specify there was a header row with the labels for the columns.

b. How many rows and columns does \texttt{rain.df} have?  How do you know?  (If there are not 5070 rows and 27 columns, you did something wrong in the first part of the problem.)
```{r columns and rows}
rowcol = c(nrow(rain.df),ncol(rain.df))
rowcol
```
I used the commands nrow and ncol, which return the number of rows of the input dataframe and number of columns, respectively. 

c. What command would you use to get the names of the columns of \texttt{rain.df}? What are those names?
```{r names of col}
colnames(rain.df)
```
The name of the ith column is V_i, where i ranges from 1 to 27. To do this, I used the function colnames(x), which returns a list of the names of the columns in the dataframe x. 

d. What command would you use to get the value at row 2, column 4? What is the value?
```{r getting 2,4}
val2 = rain.df[2,4]
val2
```
You would use the command df[i,j] to retrieve the value in the ith row and jth column. df[2,4] = 0. 

e. What command would you use to display the whole second row?  What is the content of that row?
```{r getting 2}
row2 = rain.df[2,]
row2
```
You would use the command rain.df[i,] to get the contents of the ith row. THe content of rain.df[2,] is printed above. 

f.  What does the following command do?
```{r test command}
names(rain.df) <- c("year","month","day",seq(0,23))
```
It renames the first three columns to "year", "month", and then "day", and then the next 24 columns are renamed the numbers from 0 to 23. 

g.  Create a new column called \texttt{daily}, which is the sum of the 24 hourly columns.
```{r create daily}
subset = rain.df[,4:27]
rain.df$daily = rowSums(subset)
```
h. Give the command you would use to create a histogram of the daily rainfall amounts.  Please make sure to attach your figures in your .pdf report. 
```{r create histogram}
rain.df %>% 
  ggplot(aes(x = daily)) + 
  geom_histogram() + 
  theme_bw()
```

i. Explain why that histogram above cannot possibly be right.
It cannot possibly be right because there are values in the dataset that have a negative value for total rainfall. This obviously does not make sense, as rainfall for a given day should be greater than or equal to zero. 

j. Give the command you would use to fix the data frame.
```{r fix negative}
rain.df[rain.df < 0] <- 0
```
k. Create a corrected histogram and again include it as part of your submitted report. Explain why it is
more reasonable than the previous histogram.
```{r correct histogram}
rain.df %>% 
  ggplot(aes(x = daily)) + 
  geom_histogram() + 
  theme_bw()
```
It is a much more reasonable histogram because it includes no negative values of rainfall, and because it has a distribution we would expect; specifically, right-skewed. This is because most days we would expect no rain, some days would have a little bit of rain, and a few days would have an incredible amount of rain over this period. 

***Data types***

2. (9 points, equally weighted) Make sure your answers to different parts of this problem are compatible with each other.

a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.
```
x <- c("5","12","7")
max(x)
sort(x)
sum(x)
```
b. For the next two commands, either explain their results, or why they
should produce errors.
```
y <- c("5",7,12)
y[2] + y[3]
```
c. For the next two commands, either explain their results, or why they should
produce errors.
```
z <- data.frame(z1="5",z2=7,z3=12)
z[1,2] + z[1,3]
```

3. (3 pts, equally weighted). 

a.) What is the point of reproducible code? 

b.) Given an example of why making your code reproducible is important for you to know in this class and moving forward. 

c.) On a scale of 1 (easy) -- 10 (hard), how hard was this assignment. If this assignment was hard ($>5$), please state in one sentence what you struggled with. 



